Model based reinforcement learning at FERMI FEL

### 2020-11-18 main.tex
- converging NAF2

\begin{abstract}
In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system. Both algorithms and approaches are new in this context and the main purpose of this paper is to show that reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems. In terms of sample-complexity the model-based approach is faster, while the finale performance of the model free method is superior - the so called asymptotic performance. The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency
\end{abstract}
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
\begin{itemize}
	\item Noise benchmarks with pendulum
	\item Benchmark NAF2 with pendulum
	\item What is our improvement concerning the algorithm
	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
	\item Tailored to accelerators - short horizons are standard?
	\item Improvements of the algorithms:
	\begin{itemize}
		\item Adding the noise feature
		\item Jumping feature using SAC
		\item Noise in the policy
		\item discuss different approaches to e.g. dynamic waiting time
	\end{itemize}
\end{itemize}
%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes. Another promising way is to apply reinforcement learning which has several advantages over optimisation methods:
\begin{itemize}
    \item Covers a larger class of problems.
    \item Solves the problem not always completely from the beginning as an DFO.
    \item Ability to use existing data.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
RL optimises a sequence of decisions. On critical aspect using RL is the number of iteration to train a controller.
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.

>>>>

\begin{abstract}
In this paper we discuss a model based reinforcement learning approach in comparison to a model free reinforcement learning approach applied at the FERMI FEL system. Both algorithms and approaches are new in this context and the main purpose of this paper is to show how reinforcement learning can be applied on an operational level in a feasible training time on real accelerator physics problems. In terms of sample-complexity the model-based approach is faster, while the final performance of the model free method is superior - the so called asymptotic performance. The model-based algorithm is done in a Dyna-style using an uncertainty aware model and the model-free algorithm is based on tailored deep Q-learning using some tricks to increase the sample efficiency
\end{abstract}
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
\begin{itemize}
	\item Noise benchmarks with pendulum
	\item Benchmark NAF2 with pendulum
	\item The first time model based on accelerator problem compared to a novel state of the art MFRL: NAF2
	\item Tailored to accelerators - short horizons are standard?
	\item Improvements of the algorithms:
	\begin{itemize}
		\item Adding the noise feature
		\item Jumping feature using SAC
		\item Noise in the policy
		\item discuss different approaches to e.g. dynamic waiting time
	\end{itemize}
\end{itemize}
%\tableofcontents
\section{Introduction and Motivation}
In particle accelerators one main goal is to provide stable and reproducible performance. In order to achieve this, a number of control problems have to be considered simultaneously. Especially if there is no way to model the physics, one might use optimisation techniques as e.g. derivative free optimisers (DFOs) or model-based optimisations as Gaussian processes to restore or maintain the performance. Another promising way is to apply reinforcement learning which shows several advantages over optimisation methods:
\begin{itemize}
    \item It covers a larger class of problems, RL optimises a sequence of decisions.
    \item It memorizes the problem and does not start always from scratch as an DFO.
    \item Existing data can be used.
    \item The underlying structure of the problem might be deduced.
\end{itemize}
On critical aspect using RL is the number of iteration to train a controller.
In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at Elettra Sincrotrone Trieste. Three algorithms were applied successfully showing different advantages, which will e discussed in the following.
